---
title: "Lab03: Data Cleaning with R"
author: "International Commuters"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)
```

# Overview & Goals

In this lab, you will **continue working with the web advertising bids data** to:

- Systematically **detect** and **repair** some other data quality problems using those ideas.
- Create a **cleaned version** of the dataset (e.g., `bids_clean`).
- Document your work in a **data cleaning log**.

By the end, you should have:

1. A data set with the majority of the insidious issues corrected.
2. A short narrative (or table) explaining **what you fixed** and **how**.

> *Work in the **same repository** as previous labs. Update your Jira Kanban board with new cards for this lab and keep committing your progress with clear messages.*

# Setup

## Packages

Minimally, you will need these packages:

```{r loadpackages}
library(tidyverse)
library(arrow)

library(logger)
library(glue)
library(dplyr)
library(tidyr)
library(rlang)
library(lubridate)
library(tictoc)
library(here)
library(jsonlite)
library(knitr)
library(kableExtra)
library(DT)

source(here("src", "data_cleaning.r"))
```

- `tidyverse` (especially `dplyr`, `ggplot2`, `readr`, `stringr`)
- `stringr` to modify character strings
- `lubridate` for date–time parsing
- `arrow` if you are reading/writing Parquet files

> **Task:** In a code chunk, load the packages you need.

## Load the messy bids data

Use the **same dataset** as in the earlier lab (e.g., a Parquet file in your `data/` folder).

> **Tasks:**
>
> 1. Read the messy data into R (e.g., using `arrow::read_parquet()` or `readr` functions, depending on your file).
> 2. Use `dplyr::glimpse()` and/or `summary()` to remind yourselves what variables are present and roughly what they look like.
> 3. Create a **working copy** of the data so that you can always go back to the original if needed.

You may want to keep the original as `bids_raw` and your working copy as `bids`.

```{r}
# Your code here

original_bids <- read_parquet(here("data", "bids_data_vDTR.parquet"))
bids <- original_bids %>% mutate(row_id = row_number())

expected_columns_path <- here("data", "expected_columns.csv")

if (file.exists(expected_columns_path)) {
  expected_columns <- data.frame(readr::read_csv(
    expected_columns_path,
    col_types = "ccc"
  ))
} else {
   expected_columns <- data.frame(
    column_name = c("id", "value", "timestamp"),
    expected_type = c("integer", "double", "datetime"),
    description = c("Unique ID", "Numeric Value", "Time Recorded")
  )
  readr::write_csv(expected_columns, expected_columns_path)
  message("expected_columns.csv was not found, so a default version was created.") 
}

glimpse(bids)

```


# Quick Orientation Check (10–15 minutes)

Using your prior lab work and Data Cleaning lectures:

1. In **2–4 sentences**, describe **what each row** in this dataset represents.
*Each row represents a single bid in an online advertising auction. It captures the bid's details, including the auction context, user device information, geographic data, ad specifications, and the bid's monetary and technical parameters.*

2. Identify **one or two variables** that you think should be:
   - **Mandatory** (must not be missing), and
   - Possibly a **unique key** or part of a composite key.
*Mandatory: `TIMESTAMP` = When it happened, `PRICE` = The business value*
*Unique key: As confirmed in the analysis, no natural composite key (e.g., `AUCTION_ID` + `PUBLISHER_ID`) is truly unique, as duplicate combinations exist. Therefore, the artificially added `row_id` column is necessary to serve as a unique key for each bid record.*

3. Pick **two variables** and, for each, assign at least one **data quality dimension** (e.g., completeness, validity, consistency, uniqueness, cross-field consistency).
*`PRICE` (Validity & Completeness)*
*`RESPONSE_TIME` (Validity & Consistency)*

> **Write your answers directly in this Rmd in a text section (not code).**

---

NEAL:

I already did some work on this and have a summary function that looks at the bids df and the
expected columns, as reported in the expected_columns.csv file. Can follow up with written explanation.

I already answered the unique key question in lab 2. There is no set of cols that I feel confident can be used
as a unique key. I added a row_id column to the bids dataframe to use as a unique key.

I made a slight change to the code to help it run even if the expected csv is not there, so far work 
that you've done Neal looks great!

---

```{r type summary}
# NEAL: I already did some work on this and have a summary function that looks at the bids df and the
# expected columns, as reported in the expected_columns.csv file. Can follow up with written explanation.

missing_columns <- check_columns(bids, expected_columns$column_name)
cat(glue::glue("There are {length(missing_columns)} missing column(s): \n {paste(missing_columns, collapse = ', ')}"))

bids_type_summary <- check_column_types(bids,  expected_columns)
print(bids_type_summary)

# I already answered the unique key question in lab 2. There is no set of cols that I feel confident can be used
# as a unique key. I added a row_id column to the bids dataframe to use as a unique key.

ex <- bids %>% group_by(AUCTION_ID, PUBLISHER_ID) %>%  filter(n() > 1) %>% head(5)
DT::datatable(ex,
              options = list(scrollX = TRUE, scrollY = "300px", pageLength = 5),
              class = "compact stripe") %>%
  formatStyle(
    columns = names(ex),
    backgroundColor = "white",
    color = "black"
  )

```


# Fixing Numeric & Price Issues (30–40 minutes)

Focus on the **PRICE** variable (and other numeric variables if needed).

You should investigate:

- **Type issues** (e.g., stored as character instead of numeric),
- **Impossible values** (negative values, weird “huge” values),
- **Sentinel codes** (e.g., `-999`),
- Extra non-numeric characters.

## Suggested tools

- `class()`, `unique()`, `dplyr::count()`, `summary()`
- `readr::parse_number()`
- `as.numeric()`
- `dplyr::mutate()`, `dplyr::case_when()`
- `is.na()`, `sum(is.na(...))`
- `quantile()` to inspect distributions

## Inspect PRICE

> **Tasks:**
>
> 1. Check the **type** of `PRICE` and inspect a sample of values (e.g., using `sample()` or `head()`).

```{r PRICE head}
cat(glue("The starting class (type) of the PRICE column is {class(bids$PRICE)}"))

print(head(bids$PRICE, 10))
```

> 2. In short sentences, describe what you see:
>    - Is `PRICE` numeric or character?
>    - Do you see obvious anomalies or strange values?

The 'PRICE' column is currently a character vector. There are values that appear to be numeric amounts stored as character strings. No immediate obvious anomalies like negative values or non-numeric characters in the first 10 entries, but further checks could be done with adjusted scope.


## Create a clean numeric version of PRICE

Create a **cleaned numeric version** (e.g., `PRICE_clean`) that:

- Extracts numeric content from strings (if necessary),
- Converts to numeric,
- Keeps track of values that cannot be converted (they will become `NA`).

> **Tasks:**
>
> 1. Use one or more of these:
>    - `readr::parse_number()`
>    - `stringr::str_remove_all()`
>    - `as.numeric()`
>
>    To learn about their use, type `?` before either of them (e.g., `?readr::parse_number()`)


---

NEAL:
Already done. Two issues were identified and fixed:
  1. prices are stored as character instead of numeric
  2. there are some prices that the leading 0 is an "O". The code below fixes this.

I overwrite the original column, but the original data is maintained in the original_bids dataframe.
Can change if we need to.

Code can be found in the src/data_cleaning.r file. Example usage can be found in the src/data_cleaning_example.r file.
The code is not elegant and not the easiest to modify without getting into the guts of the code. Happy to fix if needed.

---

Code

```{r PRICE conversion}
# Your code here

bids <- convert_to_numeric(bids, "PRICE", fix_leading_o = TRUE)
head(bids$PRICE, 10)
```

> 2. Count how many `NA`s you get in `PRICE_clean`. Explain why some values might be `NA` after parsing.


```{r PRICE na count}
# Your code here
na_prices <- sum(is.na(bids$PRICE))
cat(glue("There are {na_prices} NA prices"))
```


## Define and apply rules for impossible values

Now decide how to handle:

- Negative prices,
- Very large prices that are implausible,
- Sentinel codes (like `-999`).

> **Tasks:**
>
> 1. Use `dplyr::filter()` and summaries (`summary()`, `quantile()`) to identify at least **two kinds** of “bad” price values.


```{r}
# Your code here

summary(bids$PRICE)
quantile(bids$PRICE, probs = seq(0, 1, 0.01), na.rm = TRUE)
bids %>% 
  filter(PRICE < 0 | PRICE %in% c(-999, 999, 9999)) %>% 
  count(PRICE) %>% 
  arrange(desc(n))

```

-999 appears 12 times, this is a placeholder (missing value) and can be treated as NA
There are several entries that are negative, these are invalud because prices can not be negative.


> 2. Write down your **rules** for how to handle each type (e.g., set to `NA`, drop rows, cap values).

Our rules for how to handle each "bad" prices are as follow:
- values like -999, action is to set to NA, since -999 is a     common placeholder for missing or undefined values
- negative values should be set to NA since prices can't be negative and should not be treated as a real price
- implausibly high values (>10) set to na, since based on the data's distribution, prices above 10 are likely errors or outliers.
  

> 3. Implement those rules in a new variable (e.g., `PRICE_final`) using `dplyr::mutate()` and `dplyr::case_when()`.

```{r}
# Your code here

bids <- bids %>% 
  mutate(PRICE_final = case_when(
    PRICE <= 0 ~ NA_real_,
    PRICE > 10 ~ NA_real_,
    TRUE ~ PRICE))
  


```

> 4. Briefly describe your rules in text and justify them using the **validity** and **completeness** dimensions of data quality.

The following rules were applied to clean the PRICE variable:
Values less than or equal to 0, including values like -999, were set to NA because they are invalud and do not represent realistic bid prices.
Values greater than 10 are also set to NA, as they are so far outside the typical range observed (99th % < 2.1) and are implausible.
All other prices were kept as is.
These rules improve the validity of the data by removing nonsensical/placeholder entries that could bias our analysis. They also enhance completeness by clearly identifying and marking missing or unusable values as NA, rather than leaving them as misleading numerical values.


# Cleaning Categorical & String Fields (30–40 minutes)

Now work on **categorical** and **string** variables, such as:

- `DEVICE_GEO_REGION`
- `DEVICE_GEO_ZIP`
- `RESPONSE_TIME`

Remember: **this dataset only includes bids from Oregon**, so a correct region should be coded as `"OR"`. Other values in `DEVICE_GEO_REGION` represent **modified/mangled versions** of `"OR"`.

Helpful string tools:

- `stringr::str_trim()`, `stringr::str_to_upper()`, `stringr::str_replace()`
- `stringr::str_detect()`, `stringr::str_pad()`, `stringr::str_length()`
- `dplyr::count()` to see category frequencies

## DEVICE_GEO_REGION: standardizing Oregon codes

> **Tasks:**
>
> 1. Use `dplyr::count(DEVICE_GEO_REGION, sort = TRUE)` to list distinct region values and their frequencies.

```{r}
# Your code here

bids %>% 
  count(DEVICE_GEO_REGION, sort = TRUE)

```

> 2. Identify **all values** that clearly correspond to **Oregon**, just coded inconsistently (e.g., extra punctuation, different case, misspellings).

In this code I am converting everything to lowercase to make matching easier and using str_detect() to find variants


```{r}
# Your code here

bids %>% 
  mutate(region_lower = stringr::str_to_lower(stringr::str_trim(DEVICE_GEO_REGION))) %>% 
  count(region_lower, sort = TRUE) %>% 
  filter(
    str_detect(region_lower, "^or$")|
    str_detect(region_lower, "oregon") | 
    str_detect(region_lower, "xor") | 
    str_detect(region_lower, "^or[^a-z]*$")
  )

```

> 3. Define a **standard representation** (`"OR"`) and create a cleaned version (e.g., `DEVICE_GEO_REGION_clean`) where:
>    - All Oregon variants are standardized to `"OR"`.
>    - Other values (if any) are treated according to a rule you decide (e.g., set to `NA` if they cannot be interpreted as Oregon).


```{r}
# Your code here

bids <- bids %>% 
  mutate(
    DEVICE_GEO_REGION_clean = case_when(
      str_to_lower(str_trim(DEVICE_GEO_REGION)) %in% c("or", "oregon", "xor") ~ "OR",
      TRUE ~ NA_character_
    )
  )

```

> 4. In **2–3 sentences**, explain:
>    - How you detected inconsistent codes,
>    - The **set-membership rule** you are enforcing (valid region codes for this dataset).

We detected incosistent region codes by using dplyr::count() to list all distinct values of DEVICE_GEO_REGION and observed multiple variations of "Oregon" such as "Or", "oregon", and "xor". We standardized these using string manipulation functions to detect case differences and common misspellings.
The set-membership rule we enforced is that only region codes clearly representing Oregon, specifically "OR", "oregon", and "xor", are valid. All other values are treated as invalid and set to NA, since this dataset is supposed to contain only bids from Oregon.

## DEVICE_GEO_ZIP: formats and sentinels

Treat ZIP codes as **strings**, not numbers.

> **Tasks:**
>
> 1. Convert ZIP codes to character and use `dplyr::count()` to inspect common values.


```{r}
# Your code here

```

> 2. Identify suspicious ZIP values:
>    - Too short / too long,
>    - Obvious sentinel codes (e.g., `-999`, `9999`),
>    - Non-digit characters.

```{r}
# Your code here

```

> 3. Decide on a **pattern** for valid ZIP codes (e.g., Oregon 5-digit ZIPs).
>    - Consider using `stringr::str_length()`, `stringr::str_detect()` with a regular expression pattern, and `stringr::str_pad()` if necessary.

```{r}
# Your code here

```

> 4. Create a cleaned version (e.g., `DEVICE_GEO_ZIP_clean`) that:
>    - Trims whitespace,
>    - Converts sentinel codes to `NA`,
>    - Enforces your chosen ZIP pattern.

```{r}
# Your code here

```

> 5. In a short paragraph, describe your rule and why it is reasonable given the context (Oregon-only data).


## RESPONSE_TIME: prefixes and extra characters

The `RESPONSE_TIME` variable may include:

- A **text prefix** (e.g., “RESPONSE_TIME:” or typo’d variants),
- Extra characters at the end (e.g., punctuation),
- Numeric content representing a duration.

> **Tasks:**
>
> 1. Use a small `sample()` of `RESPONSE_TIME` to inspect the raw patterns.

```{r}
# Your code here

```

> 2. Write a rule for how to:
>    - Remove any “RESPONSE_TIME:” prefixes (including likely typos),
>    - Strip extra non-numeric characters from the end,
>    - Convert the remaining content to a numeric type.
> 3. Implement your rule in a cleaned variable (e.g., `RESPONSE_TIME_clean`) using `stringr::str_replace()` or related functions and `as.numeric()`.


---

Again, i am using the code from the src/data_cleaning.r file. I extract numeric values from the RESPONSE_TIME column.
I am not finding any NA's after the extraction, but have not looked for outlier values or sentinel codes.

---

```{r RESPONSE_TIME conversion}
# Your code here
bids <- convert_to_integer(bids, "RESPONSE_TIME", extract_digits = TRUE)
head(bids$RESPONSE_TIME, 10)
```

> 4. Count how many `NA`s you get in `RESPONSE_TIME_clean` and briefly interpret what those `NA`s mean.

```{r RESPONSE_TIME NA count}
# Your code here
na_response_time <- sum(is.na(bids$RESPONSE_TIME))
cat(glue("There are {na_response_time} NA response times"))
```

> 5. Connect this to the idea of **regular-expression patterns** as a tool for enforcing validity.

# Date–Time Cleaning: TIMESTAMP (20–30 minutes)

The `TIMESTAMP` field records when a bid occurred, but the **format is inconsistent** (for example, some dates may use `-` and others `/`).

Useful tools:

- `as.character()`
- `stringr::str_detect()`
- `lubridate::parse_date_time()`, `lubridate::ymd_hms()`, `lubridate::mdy_hms()`

## Inspect TIMESTAMP patterns

> **Tasks:**
>
> 1. Look at a small random sample of `TIMESTAMP` values.

```{r}
# Your code here

```

> 2. Use `stringr::str_detect()` to count how many timestamps use `"-"` versus `"/"` in the date.

---

NEAL: Did some exploring of the TIMESTAMP col but deleted most of the code once I identified the issues. The original messy code can be found at the bottom of src/data_cleaning_example.r file. Have to go back to commit 97b3c8c to see original code. Will repeat some of the relevant parts below.

TL;DR
The code below is messy and inefficient, but it was how I (NEAL) explored the TIMESTAMP and DATE cols. My conclusion, and solution, was to concatenate the date from DATE_UTC and time from TIMESTAMP as strings into an new column and then convert to POSIX.

---
```{r TIMESTAMP exploration}
# Your code here

# EXPLORATORY CODE
# ========================================================================================
# Get DATE_UTC column, which is a character initially, and convert to Date
# Get rows where complete.cases is FALSE

datetime_df <- data.frame(
  DATE_UTC_0 = original_bids$DATE_UTC,
  TIMESTAMP_0 = original_bids$TIMESTAMP
)

datetime_df$DATE <- as.Date(datetime_df$DATE_UTC_0, format = "%Y-%m-%d")

# get TIMESTAMP column, which is a character initially, split each row with delim " "
# Then convert to a dataframe and find all rows where the first column is "NA" <-- string
# Parse the TIMESTAMP column using space delimiter into date and time components
timestamp_split <- strsplit(as.character(datetime_df$TIMESTAMP_0), " ")
timestamp_df <- as.data.frame(do.call(rbind, timestamp_split), stringsAsFactors = FALSE)
colnames(timestamp_df) <- c("TIMESTAMP_date", "TIMESTAMP_time")

# Combine into one df
datetime_df <- cbind(datetime_df, timestamp_df)

DATE_date_NA_indices <- which(is.na(datetime_df$DATE))
TIMESTAMP_date_NA_indices <- which(apply(timestamp_df, 1, function(row) any(row == "NA")))

# Add new column to datetime_dfthat duplicates the TIMESTAMP_date column
# Overwrite TIMESTAMP_date2 with date_col for rows in na_row_indices
datetime_df$TIMESTAMP_date2 <- datetime_df$TIMESTAMP_date
datetime_df$TIMESTAMP_date2[TIMESTAMP_date_NA_indices] <- as.character(datetime_df$DATE[TIMESTAMP_date_NA_indices])

# Check if date_col and TIMESTAMP_date2 are equal (element-wise, after coercion if needed)
are_equal <- all(as.Date(datetime_df$TIMESTAMP_date2) == as.Date(datetime_df$DATE), na.rm = TRUE)
cat("Are date_col and TIMESTAMP_date2 equal?\n")
print(are_equal)
cat(sprintf("Number of mismatches: %d\n", sum(!are_equal, na.rm = TRUE)))


# Create a new TIMESTAMP column that is the concatenation of TIMESTAMP_date2 and TIMESTAMP_time_col
datetime_df$TIMESTAMP_corrected <- paste(datetime_df$TIMESTAMP_date2, datetime_df$TIMESTAMP_time)

# Handle multiple formats in the TIMESTAMP column
format_1 <- "%Y-%m-%d %H:%M:%S"
format_2 <- "%m/%d/%Y %H:%M:%S"

format_1_col <- as.POSIXct(datetime_df$TIMESTAMP_corrected, format = format_1, tz = "UTC")
format_2_col <- as.POSIXct(datetime_df$TIMESTAMP_corrected, format = format_2, tz = "UTC")

not_NA_format_1 <- which(!is.na(format_1_col))
not_NA_format_2 <- which(!is.na(format_2_col))

any_in_both <- any(not_NA_format_1 %in% not_NA_format_2)

combined_indices <- sort(unique(c(not_NA_format_1, not_NA_format_2)))

all_row_indices <- seq_len(nrow(datetime_df))

missing_indices <- setdiff(all_row_indices, combined_indices)

all_corrected_col <- format_1_col
all_corrected_col[not_NA_format_2] <- format_2_col[not_NA_format_2]

datetime_df$TIMESTAMP_posixct <- all_corrected_col

# Check for NA values in TIMESTAMP_posixct column
na_posixct_indices <- which(is.na(datetime_df$TIMESTAMP_posixct))
cat(sprintf("Number of NA values in TIMESTAMP_posixct: %d\n", length(na_posixct_indices)))
```

> 3. Summarize in 1–2 sentences what you observe.



## Parse TIMESTAMP into a POSIXct variable

> **Tasks:**
>
> 1. Use `lubridate::parse_date_time()` (or similar) to create `TIMESTAMP_clean`:
>    - Allow for multiple possible formats (e.g., one order for `ymd HMS` and one for `mdy HMS` to see examples scroll paste `?lubridate::parse_date_time()` in your console and scroll all the way to the bottom of the function's help tab),
>    - Specify a time zone (e.g., `"UTC"`).


---

NEAL:

Have code to clean up the TIMESTAMP col. Solution was to concatenate the date from the DATE_UTC column (consistent formatting and NA free), and the time from the TIMESTAMP col (consistent formatting and NA free). Did check to ensure all non-NA dates in TIMESTAMP col were equal to that in the DATE col.

The final code does the above concatenation and then conversion to POSIX. See convert_to_posixct() in src/data_cleaning.r for code.

Added cols to bids
TIMESTAMP_0 - original TIMESTAMP column
TIMESTAMP_date - parsed date from TIMESTAMP_0
TIMESTAMP_time - parsed time from TIMESTAMP_0
TIMESTAMP - corrected time stamp column

---
```{r }
# Your code here

bids <- convert_to_posixct(bids, "TIMESTAMP")

```

> 2. Count how many `NA`s result from parsing.

```{r}
# Your code here
TIMESTAMP_na_count <- sum(which(is.na(bids$TIMESTAMP)))
cat(glue("There are {TIMESTAMP_na_count} NAs in the TIMESTAMP column"))

```

> 3. Use `summary()` or `range()` to examine the minimum and maximum of `TIMESTAMP_clean`.

```{r}
# Your code here

```

> 4. Briefly comment on:
>    - Whether the time range seems reasonable,
>    - What kinds of entries failed to parse (if any).

# Keys, Duplicates, and Out-of-Range Values (20–30 minutes)

## Candidate keys and duplicates

Using your earlier reasoning:

> **Tasks:**
>
> 1. Choose one variable or a combination of variables that you **believe should uniquely identify** a row (a key).
> 2. Use `dplyr::count()` to check whether that key is truly unique or if there are duplicates.

```{r}
# Your code here

```

> 3. Check for **full-row duplicates** using `dplyr::distinct()` or by counting occurrences across all columns.

---

NEAL:

Already have code to check for duplicated, but need to exclude the row_id column from the check. Using duplicated() function which should only identify a row as a duplicated if it has already shown up, i.e. row of first occurrence is not listed in duplicate_indices

---
```{r}
# Your code here

duplicate_check <- find_duplicates(bids, exclude_cols = c("row_id"))
duplicate_indices <- duplicate_check$duplicate_indices
num_duplicates <- duplicate_check$num_duplicates

cat(glue("There are {num_duplicates} duplicate rows in the dataset"), "\n")
cat("first 10 duplicate rows: ",duplicate_indices[1:10])
```

> 4. Decide on a rule for handling duplicates:
>    - Drop exact duplicates?
>    - Keep first occurrence and drop the rest?
>    - Something else?
> 5. Implement your rule and create a de-duplicated dataset (e.g., `bids_nodup`).

```{r}
# Your code here

```

> 6. In 2–3 sentences, explain your choice and connect it to the **uniqueness** dimension of data quality.

## Out-of-range latitude/longitude (optional but recommended)

Look at `DEVICE_GEO_LAT` and `DEVICE_GEO_LONG`.

> **Tasks:**
>
> 1. Use `summary()` or `dplyr::summarise()` to get the minimum and maximum latitude and longitude.

```{r}
# Your code here

```

> 2. Compare the ranges to what you know about **Oregon’s** geographic coordinates (you can approximate based on general US geography; no need for exact boundaries).

```{r}
# Your code here

```

> 3. Decide which coordinates are clearly **implausible**.

```{r}
# Your code here

```

> 4. Define rules to create cleaned versions (e.g., `DEVICE_GEO_LAT_clean`, `DEVICE_GEO_LONG_clean`) where:
>    - Plausible values are kept,
>    - Implausible values are set to `NA` or corresponding rows are removed.

```{r}
# Your code here

```

> 5. Document the rules you used and why.

# Creating a Final Cleaned Dataset (10–15 minutes)

Now pull everything together.

> **Tasks:**
>
> 1. Create a final cleaned dataset (e.g., `bids_clean`) that:
>    - Uses your cleaned variables (`PRICE_final`, `DEVICE_GEO_REGION_clean`, `DEVICE_GEO_ZIP_clean`, `RESPONSE_TIME_clean`, `TIMESTAMP_clean`, cleaned lat/long, etc.),
>    - Drops unnecessary intermediate columns (e.g., `_raw`, helper strings) that you don’t need for analysis.

```{r}
# Your code here

```

> 2. Check the structure of `bids_clean` with `dplyr::glimpse()`.

```{r}
# Your code here

```

> 3. Optionally, write the cleaned dataset to disk (for your own future use) using something like `save()` (see `?save` for its use and an example) to save in `.RData` format **but do not commit the cleaned data file** to the remote repository.

```{r}
# Your code here (optional)

```

> 4. In a short paragraph, describe what your final cleaned dataset now represents and what major issues have been addressed.

# Data Cleaning Log & Reflection (10–15 minutes)

## Data cleaning log (table)

Create a succinct table summarizing the main problems and fixes.

Use the template below and fill it in with your team’s work. **One row (PRICE) is completed as an example; you should complete the remaining rows yourselves.**

| Variable | Problem Detected (brief) | Data Quality Dimension(s) | Cleaning Action (brief) |
|----------|-------------------------|---------------------------|--------------------------|
| PRICE | Includes non-numeric characters | Validity | Extract numeric portion and convert to numeric |
| DEVICE_GEO_REGION | *(team describes)* | *(team selects)* | *(team describes)* |
| DEVICE_GEO_ZIP | *(team describes)* | *(team selects)* | *(team describes)* |
| RESPONSE_TIME | *(team describes)* | *(team selects)* | *(team describes)* |
| TIMESTAMP | *(team describes)* | *(team selects)* | *(team describes)* |
| Other variable(s) | *(optional)* | *(optional)* | *(optional)* |

## Reflection (5–8 sentences)

Write a short reflection addressing:

- Which data cleaning tasks were the most **time-consuming** or **surprising**?
- Which functions or workflows (e.g., `mutate()`, `case_when()`, `str_*()`, `parse_date_time()`, `distinct()`) were most valuable?
- One thing you will do **differently** the next time you receive a new messy dataset.
- Whether/how you used generative AI for this lab and how it affected your understanding.
